---
title: "Examples"
author: "Simon Kucharsky"
date: "`r Sys.Date()`"
output: 
  rmarkdown::github_document:
    pandoc_args: --webtex
bibliography: "`r here::here('bibliography.bib')`"
csl: "`r here::here('apa.csl')`"
---

This file serves as a supplement to the article "Dynamic models of eye movements" and provides step by step explanation of the building and implementation of the modelling framework of the joint distribution of fixation locations and fixation durations.

In this folder, there are additional files that go through specific examples of the model. These examples should provide some guidance in building and implementing such a model. The models are implemented in Stan [@carpenter2017stan] in combination with the R package rstan [@team2016rstan].

# Principled Bayesian modeling

Here, we summarise several criteria and approaches we used for validating our model, following general reccomendations for Bayesian modeling, see @schad2019principled.

## Prior predictives

Each of the models is parametrized with several parameters, some of which do, and other do not have clear physical interpretation. 

## Computational faithfullness

We check whether the Bayesian inference for the implemented model is self consistent, i.e., whether the Bayesian updating of the prior leads to correct posteriors. We check that by conducting simulation-based calibration [@talts2018sbc].

## Model sensitivity

We also investigate whether and to what extend are the estimates of the paramaters biased, and to what extend are data informative to specific parameters. To do this, we plot the posterior z-scores and posterior contraction against each other for all focal parameters in the model.

## Posterior predictives


# Common features of the Stan files

All of the models we implemented share some common common computations, some of which needed to be implemented by specifying custom functions in Stan. These functions are loaded in each of the model's file by including the `load_functions.stan` file in the function block. This file loads the following files:


```{r echo=FALSE}
writeLines(readLines(here::here("stan", "helpers", "load_functions.stan")))
```

Here, we go through all of the files to document the functionality we added to Stan.

## Wald distribution

We implemented the Wald distribution in Stan.
The first function `wald_lpdf` provides the log of the density function for the Wald distribution. The density of the Wald distribution is the same as the density of the Inverse Gaussian distribution after reparametrization to parameters $\alpha$ (boundary) and $\nu$ (drift rate):

```{r echo=FALSE}
writeLines(readLines(here::here("stan", "helpers", "wald_lpdf.stan")))
```

The second function `wald_rng` implements the RNG function of a Wald distributed variable:

```{r echo=FALSE}
writeLines(readLines(here::here("stan", "helpers", "wald_rng.stan")))
```

The description of both functions can be found under Wikipedia entry for the [Inverse Gaussian distribution](https://en.wikipedia.org/wiki/Inverse_Gaussian_distribution).

## Log density of a gaussian mixture

For modelling the distribution of fixation locations, we defined the intensity function as a mixture density:

$$
\lambda(x, y) = \sum_{k=1}^{K} \pi_k \times \text{Normal}(x | \mu_{kx}, \sigma_{kx}) \times \text{Normal}(y | \mu_{ky}, \sigma_{ky}),
$$

with $\mu_{kx}$ and $\sigma_{kx}$ being the center and width of a factor $k$ in the $x$ dimension, $\mu_{yx}$ and $\sigma_{yx}$ being the center and width of a factor $k$ in the $y$ dimension, and $\pi_k$ the weight of a factor $k$.

To obtain the log of this density, we constructed the following convenience function that implements this calculation 

```{r echo=FALSE}
writeLines(readLines(here::here("stan", "helpers", "mixture_normals_lpdf.stan")))
```


The model we use relies on integration of the intensity function with the attention window. Recall that we define the intensity function as
and that we defined the attention window as a kernel of a Gaussian distribution:

$$
a(x, y | s) = \exp \left(-\frac{(x - s_x)^2}{2\sigma_a^2}\right) \times \exp \left(-\frac{(y - s_y)^2}{2\sigma_a^2}\right)
$$


```{r echo=FALSE}

```

```{r echo=FALSE}

```


## References
