---
title: "Initial Model"
author: "Simon Kucharsky"
date: "`r Sys.Date()`"
output: 
  rmarkdown::github_document:
    pandoc_args: --webtex
bibliography: ["`r here::here('bibliography.bib')`", "`r here::here('packages.bib')`"]
csl: "`r here::here('apa.csl')`"
---


Here, we will provide some more information about the initial model reported in the article. Specifically, here we explain the basic of the model's architecture, and describe additional information.

## The model

The model contains four factors (components), of which two are included in the model for *when.* The factors are the following:

1. Objects in the scene (included in the model for when)
2. Saliency of the images (included in the model for when)
3. Exploitation
4. Central bias


## Simulation

We conducted a small parameter recovery simulation to assess the model's performance. 

```{r message=FALSE, warning=FALSE, include=FALSE}
# Script with the initial model simulations: later it will be rewritten in rmarkdown for a nice github document
library(tidyverse)
library(imager)
library(plotrix)
library(gtools)
library(rstan)
library(here)
source(here::here("R", "load_image.R"))
source(here::here("R", "helpers.R"))
#source(here::here("R", "expose_helpers_stan.R"))
log_sum_exp <- matrixStats::logSumExp
overwrite_cache <- FALSE

## definition of critical radius
# subset saliency in radius of 100 from the fixation position
dist_screen <- 60  # distance from the screen (cm)
theta       <- 5   # radius of foveal vision (degrees of visual angle)
width_cm    <- 51  # width of the screen (cm)
width_pix   <- 800 # number of pixels in horizontal location
size_pix    <- width_cm / width_pix # size of one pixel
radius <- tan(theta * pi / 180)*dist_screen / size_pix


image_nr <- c(1001, 1014, 1049, 1087, 1092, 1099)
image_name  <- paste0(image_nr, ".jpg")
objects <- read.csv(here::here("data", "objects.csv")) %>% subset(image %in% image_nr)
objects_in_images <- read.csv(here::here("data", "objects_in_images.csv")) %>% subset(id_img %in% objects$id_img)
saliency <- read.csv(here::here("data", "saliency.csv")) %>% subset(image %in% image_nr)


```

The simulation is based on six images from the real data set. Below are the pictures with their downsampled saliency maps and highlighted the centers and sizes of labelled objects on the scene.

```{r}
par(mfrow = c(1, 2))
for(i in seq_along(image_name)) {
  img <- load_image(image_name[i])
  dim_img <- list(min_x = 0, min_y = 0, max_x = imager::width(img), max_y = imager::height(img))
  obj <- subset(objects, image == image_nr[i])
  sal <- subset(saliency, image == image_nr[i])
  
  plot(img, axes = FALSE)
  points(obj$x, obj$y, pch = 10, cex = 2)
   for(i in 1:nrow(obj))
     plotrix::draw.ellipse(x = obj$x[i], y = obj$y[i], a = obj$width[i]/2, b = obj$height[i]/2, angle = 0, lwd = 2)
  
  plot(imager::as.cimg(sal$value, x = max(sal$row), y = max(sal$col)), axes = FALSE)
}
par(mfrow = c(1, 1))
```

```{r}
# draw from priors
N_sim <- 20
N_ppt <- 25
N_obj <- nrow(objects)

load(here::here("documents", "initial_model_saves", "true.Rdata"))
load(here::here("documents", "initial_model_saves", "sim_data.Rdata"))
load(here::here("documents", "initial_model_saves", "fits.Rdata"))
```



```{r}
get_par <- function(par, true) {
  fit_summary <- t(sapply(fits, function(fit) summary(fit, pars = par)$summary[, c("mean", "25%", "75%"), drop=TRUE]))
  fit_summary <- as.data.frame(fit_summary)
  fit_summary$true <- true[, par]
  
  fit_summary
}

plot_par <- function(par, true) {
  df <- get_par(par, true)
  lim <- range(as.matrix(df))
  plot(df$true, df$mean, pch = 19, cex = 1, main = par, 
       xlab = "True", ylab = "Estimated", 
       xlim = lim, ylim = lim)
  segments(x0 = df$true, y0 = df$`25%`, y1 = df$`75%`)
  abline(a = 0, b = 1)
}

get_vec_par <- function(par, true) {
  fit_summary <- lapply(seq_along(fits), function(i) { 
    fit <- fits[[i]]
    out <- summary(fit, pars = par)$summary[, c("mean", "25%", "75%"), drop=TRUE]
    out <- as.data.frame(out)
    out$true <- unlist(true[i,,drop=TRUE])
    out$sim <- i
    colnames(out) <- c("est", "lower", "upper", "true", "sim")
    out
  })
  #fit_summary <- as.data.frame(fit_summary)
  
  do.call(rbind, fit_summary)
}

plot_vec_par <- function(par, true) {
  df <- get_vec_par(par, true)
  
  ggplot(df, aes(x=true, y=est, ymin=lower, ymax=upper, col=as.factor(sim))) + 
    geom_abline(intercept = 0, slope = 1) +
    geom_point() +
    geom_errorbar()
}

par(mfrow=c(2, 4))
for(par in colnames(true_parameters)) {
  plot_par(par, true_parameters)
}

par(mfrow=c(1, 4))
for(par in colnames(true_weights)) {
  plot_par(par, true_weights)
}

plot_vec_par("z_weights_obj", true_z_weights_obj)
plot_vec_par("alpha", true_alpha)
plot_vec_par("sigma_attention", true_sigma_attention)
```

