---
title: "Object oriented behavior"
author: "Simon Kucharsky"
date: "`r Sys.Date()`"
output: 
  rmarkdown::github_document:
    # pandoc_args: --webtex
bibliography: "`r here::here('bibliography.bib')`"
csl: "`r here::here('apa.csl')`"
---

```{r include=FALSE}
library(imager)
library(plotrix)
source(here::here("R", "load_image.R"))
source(here::here("R", "helpers.R"))
source(here::here("R", "expose_helpers_stan.R"))

image_nr <- 1001
image_name  <- paste0(image_nr, ".jpg")
objects <- read.csv(here::here("data", "object_familiarity", "objects.csv"))
objects <- subset(objects, image_name == image_nr)
```
This example demonstrates the implementation of a dynamic model of eye movements that assumes that only objects on the scene influence the eye movement behavior.


We will demonstrate this example on one stimulus from the article by @Xu2014beyond. Specifically, we use the following image `r image_name` available from their [public repository](https://github.com/NUS-VIP/predicting-human-gaze-beyond-pixels):

```{r echo=FALSE, fig.width=8, fig.height=6}
img <- load_image(image_name)
dim_img <- list(min_x = 0, min_y = 0, max_x = imager::width(img), max_y = imager::height(img))
plot(img, axes = FALSE)
points(objects$x, objects$y, pch = 10, cex = 2)
for(i in 1:nrow(objects))
  plotrix::draw.ellipse(x = objects$x[i], y = objects$y[i], a = objects$width[i]/2, b = objects$height[i]/2, angle = 0, lwd = 2)
```


The ellipses demarkate objects on the scene (based on their specified widths and heights).

## Model specification

Recall that we have a model for the fixation locations, specified as follows:

$$
\lambda(x, y) = \sum_{k=1}^K \pi_k \times \text{Normal}(x | \mu_{kx}, \sigma_{kx}) \times \text{Normal}(y | \mu_{ky}, \sigma_{ky}).
$$
Here our 'factors' will be the objects on the scene. We know the locations of the objects, and so we set the means $\mu$ to the centers of the objects. However, the objects can have varying size. To account the sizes of the objects, we could technically estimate the individual $\sigma$'s, but we can also use the fact that we have a data set that demarkates the width ($w$) and height ($h$) of the objects, and use a single parameter $\delta$ that stretches all the dimensions of the objects equally. The model for fixation locations is then 

$$
\lambda(x, y) = \sum_{k=1}^K \pi_k \times \text{Normal}(x | m_{kx}, \delta w_{kx}/2) \times \text{Normal}(y | m_{ky}, \delta h_{ky}/2).
$$
Our model for fixation duration is

$$
d^t \sim \text{Wald}(\alpha, \nu^t) \\
\nu^t = -\log \int \int a(x, y | s^t) \lambda(x, y) dx dy,
$$
where $s^t$ is the current fixation location.

## Implementation in Stan

The model can be translated in the Stan code as following

```{r}
writeLines(readLines(here::here("stan", "examples", "object_oriented_behavior.stan")))
#model <- rstan::stan_model(here::here("stan", "examples", "object_oriented_behavior.stan"))
```

## Prior Predictives

We created the model with the following priors:

$$
\begin{aligned}
\pi & \sim \text{Symmetric-Dirichlet}(2) \\
\delta & \sim \text{Exponential}(1) \\
\alpha & \sim \text{Normal}(2, 1)_{[0,\infty)} \\
\sigma_a & \sim \text{Gamma}(2, 0.1)
\end{aligned}
$$

```{r draw_parameters, echo=FALSE}
n_sim <- 1000
t_max <- 10
n_objects <- nrow(objects)
sim_parameters_list <- replicate(n = n_sim, expr = list(pi    = gtools::rdirichlet(1, rep(2, n_objects))[1,,drop=TRUE],
                                                        delta = rexp(1, 1),
                                                        alpha = rtnorm(1, 2, 1, 0, Inf),
                                                        sigma_attention = rgamma(1, shape = 2, rate = 0.1)
                                                        ),
                                 simplify = FALSE)
```


We simulate `r n_sim` data sets:

1) draw parameters based on their priors
2) draw 10 seconds of data or at most 100 fixations for a given set of drawn parameters


```{r draw_data, echo=FALSE}
simulate <- function(t_max, pi, delta, alpha, sigma_attention, n_max = t_max * 10){
  x <- y <- duration <- nu <- numeric()
  t <- 0
  sigma_x <- delta * objects$width / 2
  sigma_y <- delta * objects$height / 2
  
  while(t <= t_max && length(x) < n_max){
    which_factor <- sample(1:n_objects, 1, FALSE, pi)
    x_now <- rtnorm(1, objects$x[which_factor], sigma_x[which_factor], dim_img$min_x, dim_img$max_x)
    y_now <- rtnorm(1, objects$y[which_factor], sigma_y[which_factor], dim_img$min_y, dim_img$max_y)
    nu_now <- -log_integral_attention_mixture_2d(x = x_now, y = y_now, weights = pi, 
                                                 mu_x = objects$x, sigma_x = sigma_x, 
                                                 mu_y = objects$y, sigma_y = sigma_y, 
                                                 width_x = sigma_attention, width_y = sigma_attention)
    duration_now <- wald_rng(alpha, nu_now)
    x <- c(x, x_now)
    y <- c(y, y_now)
    nu <- c(nu, nu_now)
    duration <- c(duration, duration_now)
    t <- t + duration_now
  }
  
  return(data.frame(x=x, y=y, duration=duration, nu=nu))
}

sim_data_list <- lapply(seq_along(sim_parameters_list), function(i) {
  pars <- sim_parameters_list[[i]]
  d <- simulate(t_max = t_max, pi = pars$pi, delta = pars$delta, alpha = pars$alpha, sigma_attention = pars$sigma_attention, n_max = 100)
  d$sim <- i
  return(d)
  })

sim_data <- do.call(rbind, sim_data_list)
```

Below is what the model in combination with the priors predicts about the locations of fixations:

```{r pred_fixations}
plot(sim_data$x, sim_data$y, pch = 19, cex = 0.1, xlab = "x", ylab = "y", bty = "n", ylim = rev(range(sim_data$y)))
```

And below are the summaries of the distributions of fixation durations

```{r pred_durations}
par(mfrow = c(1, 3))
hist(sapply(sim_data_list, function(x) mean(x$duration)), breaks = n_sim / 50, main = "Mean fixation durations", xlab = "")
hist(sapply(sim_data_list, function(x) sd(x$duration)),   breaks = n_sim / 50, main = "SD of fixation durations", xlab = "", ylab = "")
hist(sapply(sim_data_list, nrow), breaks = n_sim / 50, main = sprintf("Number of fixations in %s sec", t_max), xlab = "", ylab = "")
```



## Computational faithfulness

We fit the model on all `r n_sim` simulated datasets.

```{r}
fits <- lapply()
```


## References